Damage localization based on vibrations

### 2016-11-16 rapport/paper_v0.tex
Add files via upload

\begin{document}
%
% paper titlesdsd
% can use linebreaks \\ within to get better formatting as desired
\title{Damage localization based on vibrations}

@@

\newtheorem{remark}{Remark}

@@

The process of implementing a damage identification strategy for aerospace,
 civil and mechanical engineering infrastructure is referred as a structural health monitoring (SHM) \cite{farrar2007introduction}.

In the most general terms, damage can be defined as changes in one system that negatively affect its current or future performance. 
In addition, a damage can only be observed by comparing two different states of the system,
 one of which is supposed to represent the initial state, and often undamaged, state.

Almost all industries want to detect damage in their infrastructure as quickly as possible whether 
 for the potential life-safety or economic impact.
Such detectetion requires these industries to perform some form of SHM.

Clearly, such damage identification has significant life-safety implications.
For instance, there are  no quantifiable methods to
 determine if buildings should be repaired after an earthquake yet.
SHM may one day provide the technology that can be used to significantly minimize
 the uncertainty associated with such post-earthquake damage assessments.
Thus additional catastrophes could be avoided.

Generally, the damage identification can be divided into four levels:
\begin{itemize}
\item detection: Is the structure damaged or not?
\item localization: Where is the damaged area located?
\item quantification: What is the extent of damage?
\item prediction: What is the remaining service life of the structure?
\end{itemize}
\vspace{2mm}

Traditional methods of damage detection based on visual inspections or experimental techniques such as radiography
 or ultrasound allow to obtain detailed information about a local damage state but requires to know area where it is located (and hope for easy access). 
These techniques may be costly, taking a long time to be performed and impractical to detect damage in large urban areas.
 Furthermore, they may fail if damage is not visibly evident. This is why other methods were invented.

(! A changer ? !) ->

The purpose of this work is the analysis of a particular data-driven damage localization method,
 namely the FRF interpolation method, and the development of a statistical extension of this method.
This paper is organized as follows. In section II we  present the existing methods (model-based and data driven methods), 
then in section III we detail the different models of vibrating structures and finally, in section IV we talk about the localization using FRF interpolation method and, finally,  its results in section V.

@@

The use of methods based on recorded responses proved to perform as valuable alternatives. 
The general idea for vibration-based damage identification is that the damage affects the physical properties (mass, damping, and stiffness) of the structure. 
And this will cause detectable changes in modal properties (natural frequencies, modal damping, and mode shapes). 
For instance, cracks cause reductions in stiffness. Therefore, by analyzing the physical properties of the structure a damage can be detected.
\\

Most of the different vibration based damage identification methods can be classified as data-driven or model-based.
 The model-based method assumes that a detailed numerical model of the structure is available for damage identification and, by vibration measurements, updates the model parameters.
 On another side, data-driven methods is based solely on the recorded response. This method is interesting for a real-time damage identification but unavailability of a numerical model often hinders the estimation of damage severity.
 A third group includes data-driven methods that use finite element (FE) information without model updating.

\subsection{Model-based methods}

In model-based methods, damage is identified through the updating of a finite element (FE) model.
 Most of the time, modal characteristics is the experimental data used for damage assessment through FE model updating.% which are extracted from measured response time histories using modal analysis techniques.
 Different types of modal data can be used. For changes in structural stiffness, model updating can be performed based only on natural frequencies or eigenvalues which are known to change in this case.
 Detailed informations about model-based methods can be found in references  %rajouter celles en-dessous
% C.-P. Fritzen, D. Jennewein, T. Kiefer, (1998). Damage detection based on model updating methods, Mechanical Systems and Signal Processing, 12(1): 163–186.
\cite{fritzen1998damage}
% Teughels A. and G. De Roeck. (2005). Damage detection and parameter identification by finite element model updating. Archives of Computational Methods in Engineering, 12(2):123-164.
\cite{teughels2005damage}
% Simoen E., De Roeck G., and Lombaert G. (2015). Dealing with uncertainty in model updating for damage assessment: a review. Mechanical Systems and Signal Processing, 56-57:123-149.
\cite{simoen2015dealing}

A REFORMULER EN DESSOUS

\subsection{Data driven methods}

Data driven method are inverse methods that use models based on experimental response data recorded on the structure instead of physical models
\cite{limongelli2016towards}.
 In these methods, the damage parameter $D$ is the variation of the damage feature $d$ between the inspection $I$ and the reference $R$ configurations: %Damage-sensitive features are extracted from data and their changes used to identify damage in the structures so the damage parameter D in these method is the variation of the damage feature f between the inspection I and the reference R configurations.
% référence TOWARDS EXTRACTION OF VIBRATION-BASED DAMAGE INDICATORS
% nom pdf : fact sheet template_vibration based_final.pdf

\begin{equation}
D=d_{I} - d_{R}
\end{equation}

Data-driven methods do not require a FE model  and may be applied with a limited number of available signals (both responses and excitations), making it an interesting method.
To extract the damage features, different signal-processing tools can be used. Therefore, depending on the tool,  data-driven methods can be classified in Fourier-based methods, Time series method and Time-variant methods.
A more detailed description of this methods can be found in references %rajouter celles en-dessous
% Fassois S.D. and Sakellariou J.S. (2007) Time-series methods for fault detection and identification in vibrating structures. Phil. Trans. R. Soc. A. 365, 411–448. doi:10.1098/rsta.2006.1929.
\cite{fassois2007time}
% Staszewski W. J.and Robertson A.N. (2007). Time–frequency and time–scale analyses for structural health monitoring. Phil. Trans. R. Soc. A. 365, 449–477. doi:10.1098/rsta.2006.1936.
\cite{staszewski2007time}.

\subsection{Combined data-driven and model-based methods}

A third group of methods for damage localization and quantification lies between the data-driven and model-based approaches.
 "This class is based on datadriven features from measurements of the reference and damaged states, which are confronted to a FE model of the structure to define damage indicators for the elements of the FE model,
 without updating its parameters" \cite{limongelli2016towards}. %rajouter référence au papier
% référence TOWARDS EXTRACTION OF VIBRATION-BASED DAMAGE INDICATORS
% nom pdf : fact sheet template_vibration based_final.pdf
% Maria Pina Limongelli, Eleni Chatzi, Michael D¨ohler, Geert Lombaert, Edwin Reynders
% https://hal.inria.fr/hal-01344178/document

@@

The discrete-time state-space model is:
\begin{equation}
\left\{
\begin{array}{ll}
x_{k+1} & = Ax_k + Bf_k \\
y_k & = Cx_k + Bf_k
\end{array}
\right.
\label{discrete}
\end{equation}
where
\begin{equation}
\begin{array}{ll}
A = exp(A_c \Delta t) \\
B= (A - I) A^{-1}_c B_c \\
C = C_c \\
D = D_c
\end{array}
\end{equation}

@@

\subsection{Obtaining acceleration measures}

We have implemented a simulation to check this method of localization. We made this on the following system: $n$ masses, connected
through springs and dampers. We denote the masses $\{m_i\}_{i=1}^n$, the springs stifness $\{k_i\}_{i=1}^n$ and the dampers $\{c_i\}_{i=1}^n$ (see Figure (\ref{springs})).

\begin{figure}
  \centering
  \includegraphics[width=0.4\textwidth]{images/ressorts.png}
  \caption{Masses connected through springs and dampers}
  \label{springs}
\end{figure}

For the simulation, we use the discrete-time state-space model (Equation (\ref{discrete})). First, we needed the matrices $M, C, K$. In this case, the matrices $M$ and $K$ are given by:
\begin{equation}
\begin{array}{l}
M = 
\begin{pmatrix} 
m_1\\ 
&\ddots\\ 
&&m_n
\end{pmatrix} 
\\
K = 
\begin{pmatrix} 
k_1 + k_2 & - k_1 \\ 
- k_1 &\ddots\\ 
& -k_i & k_i + k_{i+1} & -k_{i+1} \\
&&&\ddots& -k_n\\
&&& -k_n & k_n
\end{pmatrix} 
\end{array}
\end{equation}

As we have seen earlier, the damping matrix can't be determined by measures.
Thus, we do the same thing that in practical situations: we consider that we are in the case of Rayleigh damping: the damping 
matrix is a linear combination of the mass and stiffness matrix:
\begin{equation}
C = \alpha M + \beta K,\ \ \ \ \ \alpha, \beta \in \mathbb{R}_+
\end{equation}

With this matrices, we can calculate the matrices $A_c, B_c, C_c$ and $D_c$. To calculate the matrices $A, B$, we have to know the 
sampling period $\Delta t$. We need the frequencies of the system, because we have to respect the Nyquist–Shannon sampling theorem.
To determine the Nyquist frequency (here, it is the taller frequency of the system), we calculate the eigenvalues of $A_c$. Indeed, we have the link between the eigenvalues $\{\lambda_i\}_{i=1}^n$ and the frequencies of the system $\{f_i\}_{i=1}^n$:
\begin{equation}
\lambda_i = - \omega_i \xi_i \pm i \omega_i \sqrt{1 - \xi_i^2}
\end{equation}
where $\omega_i = 2 \pi f_i$ and $\xi_i$ is the damping ratio.

\begin{remark}
We can see that:\\
$|\lambda_i| = \sqrt{\omega_i^2 \xi_i^2 + \omega_i^2 (1 - \xi_i^2)} = \omega_i$
\end{remark}

\begin{remark}
The matrix $A_c$ has $2 \times n$ eigenvalues, but we obtain $n$ frequencies (each of them appears two times). 
\end{remark}

Now, we have the sampling period:
\begin{equation}
\Delta t = 2 \times max_{i \in \{1 ... n\} } f_i
\end{equation}

So we can calculate the matrices $A, B$ and make the simulation with the discrete-time state-space model (Equation (\ref{discrete})). To simulate the forces $f_k$, we use a white noise.

\subsection{Frequency Response Function}

The FRF correspond to the transfer function of the system. In frequency domain, if $F$ is the force that is apply on the system and $S$ is the response of this force, the FRF is given by:
\begin{equation}
H = \frac{S}{F}
\end{equation}

We want to calculate the FRF of the system, for each sensor localization. Here, we calculate an estimation of the FRF, called $H_1$ estimator. We suppose we have a vector of the forces $\{ f_k\}_{k=1}^N$ and a vector of the accelerations $\{ y_k\}_{k=1}^N$ for each sensor (the two vectors are size $N$). We need to be in frequency domain, so we calculate the discrete Fourier transform:
\begin{equation}
\begin{array}{ll}
F_n = \sum\limits_{k=1}^N f_k e^{-2i\pi kn/N} \ \ \ \ \ n \in \{1 ... N\} \\
Y_n = \sum\limits_{k=1}^N y_k e^{-2i\pi kn/N} \ \ \ \ \ n \in \{1 ... N\}
\end{array}
\end{equation}

Then we calculate the power spectra, i.e. the Auto Power spectra of the inputs and the outputs
and the Cross Power spectra between the inputs and outputs respectively given by:

\begin{equation}
\begin{array}{ll}
S_{ff}(n) = F_n F_n^H \ \ \ \ \ n \in \{1 ... N\} \\
S_{yf}(n) = Y_n F_n^H \ \ \ \ \ n \in \{1 ... N\}
\end{array}
\end{equation}
where $M^H$ is the Hermitian of the matrix $M$. Thus, the FRF is given by:
\begin{equation}
H_1(n) = S_{yf}(n) S_{ff}(n)^{-1}
\end{equation}

\begin{remark}
If we consider each sensor separately, $F_n$ and $Y_n$ are scalars, so we have:
\begin{equation}
\begin{array}{lll}
H_1(n) & = & S_{yf}(n) S_{ff}(n)^{-1} \\
& = & Y_n F_n^H (F_n F_n^H)^{-1} \\
& = & Y_n / F_n
\end{array}
\end{equation}
\end{remark}

\subsection{Interpolation}

\subsection{Result}

>>>>

\begin{document}
%
% paper title
% can use linebreaks \\ within to get better formatting as desired
\title{Damage localization based on vibrations}

@@

@@

There are four levels for damage identification:
\begin{itemize}
\item detection: Is the structure damaged or not?
\item localization: Where is the damaged area located?
\item quantification: What is the extent of damage?
\item prediction: What is the remaining service life of the structure?
\end{itemize}

@@

@@

The discret-time state-space model is:
\begin{equation}
\left\{
\begin{array}{ll}
x_{k+1} & = Ax_k + Bf_k \\
y_k & = Cx_k + Bf_k
\end{array}
\right.
\end{equation}
where
\begin{equation}
\begin{array}{ll}
A = exp(A_c \Delta t) \\
B= (A - I) A^{-1}_c B_c \\
C = C_c \\
D = D_c
\end{array}
\end{equation}

@@

%Thus, we consider that we are in the case of Rayleigh damping: the damping 
%matrix is a linear combination of the mass and stiffness matrix:
